### 2021-04-26 paper/natastron/ArXiV/paper.tex
updated arxiv

\textbf{Studies using asteroseismic ages and rotation rates from star-spot rotation have indicated that standard age-rotation relations may break down roughly half-way through the main sequence lifetime, a phenomenon referred to as weakened magnetic braking. While rotation rates from spots can be difficult to determine for older, less active stars, rotational splitting of asteroseismic oscillation frequencies can provide rotation rates for both active and quiescent stars, and so can confirm whether this effect really takes place on the main sequence.\\
We obtained asteroseismic rotation rates of 91 main sequence stars showing high signal-to-noise modes of oscillation.
Using these new rotation rates, along with effective temperatures, metallicities and seismic masses and ages, we built a hierarchical Bayesian mixture model to determine whether the ensemble more closely agreed with a standard rotational evolution scenario, or one where weakened magnetic braking takes place. The weakened magnetic braking scenario was found to be $98.4\%$ more likely for our stellar ensemble, adding to the growing body of evidence for this stage of stellar rotational evolution. This work represents the largest catalogue of seismic rotation on the main sequence to date, opening up possibilities for more detailed ensemble analysis of rotational evolution with \textit{Kepler}.}

@@

\textbf{Asteroseismic main sequence targets:} In order to obtain robust asteroseismic rotation rates, we required detections of multiples of both dipole (denoted as $\ell = 1$) and quadrupole ($\ell = 2$) p mode (\textit{Pressure} modes, the type of oscillation visible in main sequence stars) oscillations for each star, of which the latter have significantly lower signal-to noise. Radial oscillations ($\ell = 0$) are also visible, but do not split. In this work, we studied a sample of 95 of the highest signal-to-noise targets observed by \textit{Kepler}, combining the `Kages' \cite{silvaaguirre+2015,davies+2016} and LEGACY \cite{lund+2017, silvaaguirre+2017} catalogues. For these catalogues the mode extraction through frequency fitting \cite{davies+2016, lund+2017} and modelling using mode frequencies to obtain stellar parameters \cite{silvaaguirre+2015, silvaaguirre+2017} are covered in separate papers. Further details on these data can be found in \textbf{Methods}.

@@

We compared our results with both published and unpublished asteroseismic rotation rates \cite{nielsen+2015,davies+2015,davies+2016, lund+2017, benomar+2018}, and found that the three stars with the lowest angles of inclination ($< 10^\circ$) were inconsistent with multiple other studies. For stars viewed near pole-on the split modes of oscillation have extremely low power, making them hard to detect reliably \cite{lund+2014}. We flagged our rotation rates for these stars, and did not use them in the next steps in our analysis, leaving a sample of 91 stars for which we considered our asteroseismic rotation measurements to be robust (for a full justification, see \textbf{Supplementary Information}).\\

@@

\textbf{Conclusions:} In this work, we obtained asteroseismic rotation rates for 91 stars using a hierarchical Bayesian model to fit for oscillation frequencies, convective background, and rotational splitting simultaneously. We validated these results against previously published and unpublished asteroseismic rotation measurements, making our new rotation catalogue the largest self-consistent sample of asteroseismic rotation in main sequence stars to date. We evaluated the stars in our sample against synthetic population \rtwo{distribution} models of the \textit{Kepler} field evolved under either a `standard' model of rotational evolution, or one where weakened magnetic braking takes place. This evaluation was done using a Bayesian mixture model to determine whether the ensemble preferred one model over another, and was done simultaneously in asteroseismic rotation, mass, age, and spectroscopic temperature and metallicity.\\

>>>>

\textbf{Studies using asteroseismic ages and rotation rates from star-spot rotation have indicated that standard age-rotation relations may break down roughly half-way through the main sequence lifetime, a phenomenon referred to as weakened magnetic braking. While rotation rates from spots can be difficult to determine for older, less active stars, rotational splitting of asteroseismic oscillation frequencies can provide rotation rates for both active and quiescent stars, and so can confirm whether this effect really takes place on the main sequence.\\
We obtained asteroseismic rotation rates of 91 main sequence stars showing high signal-to-noise modes of oscillation.
Using these new rotation rates, along with effective temperatures, metallicities and seismic masses and ages, we built a hierarchical Bayesian mixture model to determine whether the ensemble more closely agreed with a standard rotational evolution scenario, or one where weakened magnetic braking takes place. The weakened magnetic braking scenario was found to be $98.4\%$ more likely for our stellar ensemble, adding to the growing body of evidence for this stage of stellar rotational evolution. This work represents a large catalogue of seismic rotation on the main sequence, opening up possibilities for more detailed ensemble analysis of rotational evolution with \textit{Kepler}.}

@@

\textbf{Asteroseismic main sequence targets:} In order to obtain robust asteroseismic rotation rates for solar-like oscillators, we required detections of multiples of both dipole (denoted as $\ell = 1$) and quadrupole ($\ell = 2$) p mode (\textit{Pressure} modes, the type of oscillation visible in main sequence stars) oscillations for each star, of which the latter have significantly lower signal-to noise. Radial oscillations ($\ell = 0$) are also visible, but do not split. In this work, we studied a sample of 95 of the highest signal-to-noise targets observed by \textit{Kepler}, combining the `Kages' \cite{silvaaguirre+2015,davies+2016} and LEGACY \cite{lund+2017, silvaaguirre+2017} catalogues. For these catalogues the mode extraction through frequency fitting \cite{davies+2016, lund+2017} and modelling using mode frequencies to obtain stellar parameters \cite{silvaaguirre+2015, silvaaguirre+2017} are covered in separate papers. Further details on these data can be found in \textbf{Methods}.

@@

We compared our results with both published and unpublished asteroseismic rotation rates for stars in our sample \cite{nielsen+2015,davies+2015,davies+2016, lund+2017, benomar+2018}, and found that the three stars with the lowest angles of inclination ($< 10^\circ$) were inconsistent with multiple other studies. For stars viewed near pole-on the split modes of oscillation have extremely low power, making them hard to detect reliably \cite{lund+2014}. We flagged our rotation rates for these stars, and did not use them in the next steps in our analysis, leaving a sample of 91 stars for which we considered our asteroseismic rotation measurements to be robust (for a full justification, see \textbf{Supplementary Information}).\\

@@

\textbf{Conclusions:} In this work, we obtained asteroseismic rotation rates for 91 stars using a hierarchical Bayesian model to fit for oscillation frequencies, convective background, and rotational splitting simultaneously. We validated these results against previously published and unpublished asteroseismic rotation measurements, producing a catalogue with a large self-consistent sample of asteroseismic rotation in main sequence stars. We evaluated the stars in our sample against synthetic population \rtwo{distribution} models of the \textit{Kepler} field evolved under either a `standard' model of rotational evolution, or one where weakened magnetic braking takes place. This evaluation was done using a Bayesian mixture model to determine whether the ensemble preferred one model over another, and was done simultaneously in asteroseismic rotation, mass, age, and spectroscopic temperature and metallicity.\\
